### 2004-02-09 resumo.tex

\noindent
Resumo da Tese apresentada à COPPE/UFRJ como parte dos requisitos necessários
para a obtenção do grau de Doutor em Ciências (D.Sc.)

\vspace{1.5cm}

\begin{center}
TÍTULO DA TESE
\vspace{1cm}

André Rabello dos Anjos
\vspace{1cm}

Julho/2004
\end{center}
\vspace{1.5cm}

\noindent
Orientador: José Manoel de Seixas
\vspace{1.5cm}

\noindent
Programa: Engenharia Elétrica
\vspace{2cm}

Blah blah blah blah blah blah blah blah blah blah blah blah blah blah blah
blah blah blah blah blah blah blah blah blah blah blah blah blah. Blah blah
blah blah blah blah blah blah blah blah blah blah blah blah blah blah blah
blah blah blah blah blah blah blah blah blah blah. Blah blah blah blah blah
blah blah blah blah blah blah blah blah blah blah blah blah blah blah blah
blah blah blah blah blah blah blah blah. Blah blah blah blah blah blah blah
blah blah blah blah blah blah blah blah blah blah blah blah blah blah blah
blah blah blah blah blah.

\clearpage

\noindent
Abstract of Thesis presented to COPPE/UFRJ as a partial fulfillment of the
\linebreak requirements for the degree of Doctor of Science (D.Sc.)

\vspace{1.5cm}

\begin{center}
THESIS TITLE
\vspace{1cm}

André Rabello dos Anjos
\vspace{1cm}

July/2004
\end{center}
\vspace{2cm}

\noindent
Advisor: José Manoel de Seixas
\vspace{2cm}

\noindent
Department: Electrical Enginnering
\vspace{2cm}

Blah blah blah blah blah blah blah blah blah blah blah blah blah blah blah
blah blah blah blah blah blah blah blah blah blah blah blah blah. Blah blah
blah blah blah blah blah blah blah blah blah blah blah blah blah blah blah
blah blah blah blah blah blah blah blah blah blah. Blah blah blah blah blah
blah blah blah blah blah blah blah blah blah blah blah blah blah blah blah
blah blah blah blah blah blah blah blah. Blah blah blah blah blah blah blah
blah blah blah blah blah blah blah blah blah blah blah blah blah blah blah
blah blah blah blah blah.