### 2004-02-13 atlas.tex
Added new chapter on ATLAS
git-svn-id: file:///Users/andre/svn/andre_phd/thesis/trunk@4 b2e9302d-b4d2-0310-80c4-8586bce995ec

%% Hello emacs, this is -*- latex -*-
\typeout{ ====================================================================}
\typeout{ This is file atlas.tex, created at 13-Feb-2004 }
\typeout{ Maintained by Andre Rabello dos Anjos <Andre.dos.Anjos@cern.ch> }
\typeout{ ====================================================================}

\chapter{O Experimento ATLAS}
\label{chap:atlas}